Dynamical modeling of tidal streams

### 2013-12-28 tex/ms.tex
streams: re-arrange figures
streams: tweak
streams: spelling

First, we compute the distribution of $\Delta \opar$, the frequency
offset between stream members and the progenitor, given the angle
offset $\Delta \apar$ as follows
\begin{equation}\label{eq:poparapargeneral}
\begin{split}
  p(\Delta \opar|\Delta \apar) & = \int \dd \Delta t\,p(\Delta \opar,\Delta t|\Delta \apar)\, \\
  & = \int \dd \Delta t \,p(\Delta \apar | \Delta \opar,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
  & \approx \int \dd \Delta t \,\delta(\Delta \apar - \Delta \opar\,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
  & = \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar,\frac{\Delta \apar}{\Delta \opar})\,,
\end{split}
\end{equation}
where in the penultimate step we have approximated the initial angle
distribution as a delta function (as the initial angle offset is small
with respect to the final offset). This is a fully general
expression. For the fiducial stream model, this can be simplified to
\begin{equation}\label{eq:poparapar}
\begin{split}
  p(\Delta \opar|\Delta \apar) \propto &\ \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar)\,,\qquad 0 < \frac{\Delta \apar}{\Delta \opar} < t_d\,.\\
  & 0\,,\qquad \mathrm{otherwise}\,.
\end{split}
\end{equation}
\equationname~(\ref{eq:poparapar}) shows why we chose to model
$p(\Delta \opar)$ as a Gaussian multiplied with $|\Delta
\opar|$. $p(\Delta \opar|\Delta \apar)$ is then a Gaussian with mean
$\Delta \Omega^m$ and variance $\sigma^2_{\Omega,1}$, for $\Delta
\opar > \Delta \apar / t_d$, and zero otherwise. The mean and variance
of such a Gaussian is straightforward to calculate in terms of error
functions. The top left panel of \figurename~\ref{fig:gd1_apar} shows
the distribution of $|\Delta \opar|$ versus $|\Delta \apar|$ (absolute
value in order to show the trailing and leading stream together) for
the simulated stream, as well as the mean and dispersion calculated
from the fiducial model. The dashed line shows $\Delta \apar = \Delta
\Omega^m \,t_d$, approximately the angle where the finite disruption
time starts to influence the mean $\Delta \opar$ of the stream. The
finite disruption time, $\approx 4.5\Gyr$ in this case, of a tidal
stream means that for stream members to have reached very large angle
differences with respect to the progenitor, they must have been
removed at large frequency differences. Around $\Delta \Omega^m
\,t_d$, average stream stars, \ie, those removed with the average
frequency offset, do not have a large enough offset to reach large
angle offsets. Therefore, even though the average frequency offset
does not change over time (see \figurename~\ref{fig:gd1_times}), the
average frequency offset, and therefore the average orbit, does change
with distance from the progenitor. It is in this sense that streams do
not follow a single orbit. This is discussed further in BOVY:REF.

@@

@@

@@

\begin{figure}
  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
  \caption{Illustration of the determination of the average stream
    track described in \sectionname~\ref{sec:track}. The small black
    dots in each panel are the points along the progenitor track at
    which the Jacobian $\partial (\veco,\veca) / \partial
    (\vecx,\vecv)$ is calculated. The large black dots are the
    coresponding points on the stream where the track is
    calculated. The dashed line in each panel is the progenitor's
    orbit, while the solid line and dash-dotted lines are the mean
    track and $2\sigma$ dispersion.}\label{fig:track}
%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
\end{figure}

\begin{figure}[t!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
    longitude. The top panel shows the stream in Galactic latitude and
    the bottom panel shows the distance from the Sun. The progenitor
    is shown as the large gray dot and its orbit as the dashed
    lines. The predicted track of the stream is shown as the solid
    line. Insets show various zoomed-in parts of the stream to show
    the stream, progenitor orbit, and predicted stream track up
    close.}\label{fig:gd1_lbd}
%python plot_stream.py ../tex/gd1_evol_lb.ps
%python plot_stream.py ../tex/gd1_evol_ld.ps
\end{figure}
\begin{figure}[t!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
  \caption{Simulated stream velocities as a function of Galactic
    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
    (top panel), proper motion in the direction of Galactic longitude
    (middle panel), and in the direction of Galactic latitude (bottom
    panel). Insets again show various interesting parts of the stream
    up close.}\label{fig:gd1_lbv}
%python plot_stream.py ../tex/gd1_evol_lvlos.ps
%python plot_stream.py ../tex/gd1_evol_lpmll.ps
%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
\end{figure}

>>>>

First, we compute the distribution of $\Delta \opar$, the frequency
offset between stream members and the progenitor, given the angle
offset $\Delta \apar$ as follows
\begin{equation}\label{eq:poparapargeneral}
\begin{split}
  p(\Delta \opar|\Delta \apar) & = \int \dd \Delta t\,p(\Delta \opar,\Delta t|\Delta \apar)\, \\
  & = \int \dd \Delta t \,p(\Delta \apar | \Delta \opar,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
  & \approx \int \dd \Delta t \,\delta(\Delta \apar - \Delta \opar\,\Delta t)\,p(\Delta \opar,\Delta t)\,,\\
  & = \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar,\frac{\Delta \apar}{\Delta \opar})\,,
\end{split}
\end{equation}
where in the penultimate step we have approximated the initial angle
distribution as a delta function (as the initial angle offset is small
with respect to the final offset). This is a fully general
expression. For the fiducial stream model, this can be simplified to
\begin{equation}\label{eq:poparapar}
\begin{split}
  p(\Delta \opar|\Delta \apar) \propto &\ \frac{1}{\left|\Delta \opar\right|}\,p(\Delta \opar)\,,\qquad 0 < \frac{\Delta \apar}{\Delta \opar} < t_d\,.\\
  & 0\,,\qquad \mathrm{otherwise}\,.
\end{split}
\end{equation}
\Eqnname~(\ref{eq:poparapar}) shows why we chose to model
$p(\Delta \opar)$ as a Gaussian multiplied with $|\Delta
\opar|$. $p(\Delta \opar|\Delta \apar)$ is then a Gaussian with mean
$\Delta \Omega^m$ and variance $\sigma^2_{\Omega,1}$, for $\Delta
\opar > \Delta \apar / t_d$, and zero otherwise. The mean and variance
of such a Gaussian is straightforward to calculate in terms of error
functions. The top left panel of \figurename~\ref{fig:gd1_apar} shows
the distribution of $|\Delta \opar|$ versus $|\Delta \apar|$ (absolute
value in order to show the trailing and leading stream together) for
the simulated stream, as well as the mean and dispersion calculated
from the fiducial model. The dashed line shows $\Delta \apar = \Delta
\Omega^m \,t_d$, approximately the angle where the finite disruption
time starts to influence the mean $\Delta \opar$ of the stream. The
finite disruption time, $\approx 4.5\Gyr$ in this case, of a tidal
stream means that for stream members to have reached very large angle
differences with respect to the progenitor, they must have been
removed at large frequency differences. Around $\Delta \Omega^m
\,t_d$, average stream stars, \ie, those removed with the average
frequency offset, do not have a large enough offset to reach large
angle offsets. Therefore, even though the average frequency offset
does not change over time (see \figurename~\ref{fig:gd1_times}), the
average frequency offset, and therefore the average orbit, does change
with distance from the progenitor. It is in this sense that streams do
not follow a single orbit. This is discussed further in BOVY:REF.

@@

\begin{figure}
  \includegraphics[width=0.8\textwidth,clip=]{track_aa.ps}\\
  \includegraphics[width=0.8\textwidth,clip=]{track_xz.ps}\\
  \includegraphics[width=0.8\textwidth,clip=]{track_ld.ps}
  \caption{Illustration of the determination of the average stream
    track described in \sectionname~\ref{sec:track}. The small black
    dots in each panel are the points along the progenitor track at
    which the Jacobian $\partial (\veco,\veca) / \partial
    (\vecx,\vecv)$ is calculated. The large black dots are the
    corresponding points on the stream where the track is
    calculated. The dashed line in each panel is the progenitor's
    orbit, while the solid line and dash-dotted lines are the mean
    track and $2\sigma$ dispersion.}\label{fig:track}
%python illustrate_track.py ../tex/track_aa.ps ../tex/track_xz.ps ../tex/track_ld.ps
\end{figure}

@@

\begin{figure}[t!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lb.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_ld.ps}
  \caption{Simulated stream after $5.011\Gyr$ as a function of Galactic
    longitude. The top panel shows the stream in Galactic latitude and
    the bottom panel shows the distance from the Sun. The progenitor
    is shown as the large gray dot and its orbit as the dashed
    lines. The predicted track of the stream is shown as the solid
    line. Insets show various zoomed-in parts of the stream to show
    the stream, progenitor orbit, and predicted stream track up
    close.}\label{fig:gd1_lbd}
%python plot_stream.py ../tex/gd1_evol_lb.ps
%python plot_stream.py ../tex/gd1_evol_ld.ps
\end{figure}

\begin{figure}[tp!!]
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lvlos.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmll.ps}
  \includegraphics[width=\textwidth,clip=]{gd1_evol_lpmbb.ps}
  \caption{Simulated stream velocities as a function of Galactic
    longitude. Same as \figurename~\ref{fig:gd1_lbd}, but for the
    stream in heliocentric line-of-sight velocity $V_{\mathrm{los}}$
    (top panel), proper motion in the direction of Galactic longitude
    (middle panel), and in the direction of Galactic latitude (bottom
    panel). Insets again show various interesting parts of the stream
    up close.}\label{fig:gd1_lbv}
%python plot_stream.py ../tex/gd1_evol_lvlos.ps
%python plot_stream.py ../tex/gd1_evol_lpmll.ps
%python plot_stream.py ../tex/gd1_evol_lpmbb.ps
\end{figure}

@@
